---
title: "Week 4 Data Analysis Script"
author: "Miles Anim-Appiah"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape2)
```


## Overview

This report presents preliminary data analysis for the week 3 example data. Specifically, this will report basic decriptive statistics as well as report tests of the following predictions:

1. People will feel more guilt when:
  + 1.1 They caused the shock (vs. their partner)
  + 1.2 Their partner was morally good (vs. bad)
  + 1.3 The shock was stronger (high > medium > low)
2. Women will report more guilt than men
3. Agreeable people will report more guilt
4. Neurotic people will report more guilt

and

5. People will be more generous when:
  + 5.1 They caused the shock (vs. theur partner)
  + 5.2 The partner was morally good (vs. bad)
  + 5.3 The shock was stronger (high > medium > low)
6. Women will be more generous than men
7. Agreeable people will be more generous
8. Neurotic people will be more generous

```{r load}

#Load the data
data<-read.csv("week 3/processed data/week 3 processed data.csv")

#Recode variables
data$PIN<-as.factor(data$PIN)
data$sex<-as.factor(data$sex)
data$shock<-as.factor(data$shock)
data$shockCause<-as.factor(data$shockCause)
data$pMoral<-as.factor(data$pMoral)

```

## Descriptives

Descriptive statistics for guilt
```{r descGuilt}

c(summary(data$guilt),"SD"=sd(data$guilt,na.rm=T))

```

Descriptive statistics for generosity
```{r descGenerosity}

#Generosity before the shock
c(summary(data$preShare),"SD"=sd(data$preShare,na.rm=T))

#Generosity after the shock
c(summary(data$postShare),"SD"=sd(data$postShare,na.rm=T))

```

Descriptive statistics for personality
```{r}
#Descriptives for agreeableness
c(summary(data$agree),"SD"=sd(data$agree,na.rm=T))

#Decriptives for neuroticism
c(summary(data$neuro),"SD"=sd(data$neuro,na.rm=T))

```

## Guilt

### Prediction 1.1

Prediction 1.1 was supported. People felt more guilty when they caused the shock than when their partner did.
```{r p1.1}
#Run the t-test for prediction 1.1
p1.1<-t.test(data$guilt~data$shockCause)

#Output
p1.1
```

### Prediction 1.2
Prediction 1.2 was supported. People felt more guilty when their partner was morally good.
```{r p1.2}
#Run the t-test for prediction 1.2
p1.2<-t.test(data$guilt~data$pMoral)

#Output
p1.2
```

### Prediction 1.3
Prediction 1.3 was supported. People felt more guilty when their partner received a stronger shock.
```{r 1.3}
#Run the one-way ANOVA for prediction 1.3
p1.3<-lm(guilt~shock,data=data)

#Output
anova(p1.3)
tapply(data$guilt,data$shock,function(x) mean(x,na.rm=T))
```

### Prediction 2
Prediction 2 was not supported. Women did not report more guilt than men.
```{r p2}
#Run the T-test for prediction 2
p2<-t.test(data$guilt~data$sex)

#output
p2
```

### Prediction 3
Prediction 3 was supported. Agreeable people were more guilty.
```{r p3}
#Run the correlation test for prediction 3 
p3<-cor.test(data$guilt,data$agree)

#Output
p3
```

### Prediction 4
Prediction 4 is not supported. Neurotic people were not more guilty.
```{r p4}
#Run the correlation test for prediction 4
p4<-cor.test(data$guilt,data$neuro)

#Output
p4
```

## Generosity

```{r melt}
dataMelt<-melt(data,
               measure.vars=c("preShare", "postShare"),
               variable.name=c("prePost"),
               value.name=c("share")
               )
```

### Prediction 5.1
Prediction 5.1 was supported. People were more generous when they caused the shock and were actually less generous when the partner caused the shock
```{r p5.1}
#Run the within subjects ANOVA for Prediction 5.1
p5.1<-aov(share~prePost*shockCause+Error(PIN),data=dataMelt)

#Output
summary(p5.1)
tapply(dataMelt$share,dataMelt[c(24,21)],function(x) mean(x,na.rm=T))
```

